On the number of tiles touched by a\\ line segment on a rectangular grid

### 2021-08-02 paper.tex
Added figures for proof of theorem of average number

% *** Prepare and include two subfigures
Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**. The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$. It is easy to see that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$. The latter two events are exclusive, and by symmetry have the same probability. Therefore
\begin{equation}
\label{eq: Pr i = 2 Pr}
\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
\end{equation}

Consider the event $x_2 \geq a(n-1)$. There are three possibilities depending on $x_1$, $n$ and $\len$. If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$. This is depicted in Figure **a**. If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$, as seen in Figure **b**. Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$. Note that the value $y_1$ is irrelevant for this.

>>>>

Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure~\ref{fig: Pr_i}. The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$. It is clear from the figure that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$. The latter two events are exclusive, and by symmetry have the same probability. Therefore
\begin{equation}
\label{eq: Pr i = 2 Pr}
\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
\end{equation}

\begin{figure}
\centering%
\subfigure[$i$ can be equal or greater than $n$ for all $x_1$, $0 \leq x_1 \leq a$]{%
\label{fig: Pr_i_full}%
\includegraphics[width=.48\textwidth]{Pr_i_full_1p35}%
}\hfill%
\subfigure[$i$ can be equal or greater than $n$ only for $a(n-1)-\len \leq x_1 \leq a$]{%
\label{fig: Pr_i_notfull}%
\includegraphics[width=.48\textwidth]{Pr_i_notfull_1p35}%
}%
\caption{Conditions for the normalized width of the canonical rectangle, $i$, being equal or greater than a given $n$. Example with $a=1.35$, $b=1$
}%
\label{fig: Pr_i}
\end{figure}%

Consider the event $x_2 \geq a(n-1)$. There are three possibilities depending on $x_1$, $n$ and $\len$. If $a(n-1) < \len$, the length $\len$ is enough for $x_2$ to exceed $a(n-1)$ for some angles $\orient$, regardless of $x_1$. This is depicted in Figure~\ref{fig: Pr_i_full}, where the thicker line on the arc represents those angles for which $x_2 > a(n-1)$, for a specific $x_1$. If $\len \leq a(n-1) < \len+a$, the length is enough only if $x_1 \geq a(n-1)-\len$. This can be seen in Figure~\ref{fig: Pr_i_notfull}, where the thicker line again represents the angles for which this happens, for a specific $x_1 \geq a(n-1)-\len$. Lastly, if $\len+a \leq a(n-1)$ it is impossible that $x_2$ exceeds $a(n-1)$. Note that the value $y_1$ is irrelevant for this.
